\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\graphicspath{{figures/}}
% for color box
\usepackage{tcolorbox}
\tcbuselibrary{listings, skins, breakable}
% math
\usepackage{amsmath, amssymb, amsthm}
% for float formatting options
\usepackage{float}
% tikz
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning, shapes.geometric, fit}
% for tikz formatting
\usepackage{adjustbox}
% for links
\usepackage[colorlinks=true, linkcolor=blue, urlcolor=magenta,
citecolor=violet]{hyperref}
% for glossary
\usepackage[nonumberlist]{glossaries}
\makeglossaries
% for table
\usepackage{booktabs}
\usepackage{tabularx}
% for fancy verbtim
\usepackage{fancyvrb}

\title{block-game}
\author{Joseph Grimsic}
\date{\today}
\usepackage[letterpaper, top=1.3in, bottom=1.3in, left=1.3in,
right=1.3in, heightrounded]{geometry}
\renewcommand{\baselinestretch}{1.15} % Line spacing
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.8em}

\begin{document}
\maketitle

% \tableofcontents
% \newpage

\section*{Representing the world}
\paragraph*{how to: represent as meshes?}
represent the world as meshes that we can update per frame

\paragraph*{chunk updates}
the solution is to separate the meshes into chunks. when a chunk is
updated, we update the corresponding mesh, as opposed to updating the
whole world as one mesh

\paragraph*{chunk size?}
maybe 16 voxels wide as tribute to minecraft. we could probably optimize this

\paragraph*{how to: represent individual voxels?}
do we want a bitmap? like, 2/3d array of block id's?

\paragraph*{how to: track chunk delta? }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

\section*{Updates}

\subsection*{bomb your friends}
\begin{itemize}
  \item talk to/pay an npc, and select a player to bomb/nuke to call
    in ordnance.
  \item explosive clay/compound generates in the terrain.
  \item tiers of explosives.
\end{itemize}

\subsection*{peace and love}
\begin{itemize}
  \item do something nice for your friends.
\end{itemize}

\subsection*{item rendering}
\begin{itemize}
  \item two kinds of items:
    \begin{itemize}
      \item blocks, rendered in 3d.
      \item items, rendered in 2d.
    \end{itemize}
  \item items rotate when dropped on the ground.
\end{itemize}

\subsection*{friends update}
\begin{itemize}
  \item host and join servers.
  \item game chat.
  \item proximity chat?
  \item telephone?
  \item shove friend into a ragdoll state?
  \item netcode everything?
\end{itemize}

\subsection*{water update?}
\begin{itemize}
  \item swim mechanics.
  \item lake/ocean generation.
  \item dynamic water?
  \item other liquids?
\end{itemize}

\subsection*{logistics: netcode}
\begin{itemize}
  \item for netcode, where does the validation happen for ``actions''?
  \item fight against hackers?
  \item does every ``action'' need to be requested by the client?
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

\section*{Multiplayer prototype (client--server \& offline)}

This section documents the GDScript multiplayer prototype added to the
project: a small client--server netcode layer and an optional UI so the
same game can run online (host/join) or offline without separate code
paths.

\subsection*{Files and autoload}

\begin{itemize}
  \item \texttt{network\_manager.gd}: core networking logic. Registered
    as an autoload named \texttt{NetworkManager} in
    \texttt{project.godot}, so any scene can call its functions.
  \item \texttt{multiplayer\_menu.gd}: UI script for host/join/leave and
    offline.
  \item \texttt{multiplayer\_menu.tscn}: scene with buttons (Host, Join,
    Leave, Offline) and optional IP/port fields and status label. Attach
    the script to a \texttt{Control} that has child nodes with the
    names \texttt{HostButton}, \texttt{JoinButton}, \texttt{LeaveButton},
    \texttt{OfflineButton}, and optionally \texttt{IPEdit},
    \texttt{PortEdit}, \texttt{StatusLabel}.
\end{itemize}

\subsection*{NetworkManager API (for developers)}

\paragraph*{Starting and stopping sessions}
\begin{itemize}
  \item \texttt{host\_game(port = 8080)}: creates an
    \texttt{ENetMultiplayerPeer} server on the given port and assigns it
    to \texttt{multiplayer.multiplayer\_peer}. Returns \texttt{true} on
    success. Emits \texttt{server\_started}.
  \item \texttt{join\_game(ip = "127.0.0.1", port = 8080)}: creates a
    client peer and connects to the given address. Returns
    \texttt{true} on success. Emits \texttt{connection\_succeeded} or
    \texttt{connection\_failed}.
  \item \texttt{leave\_game()}: closes the current peer and clears it;
    use when returning to menu or disconnecting.
  \item \texttt{start\_offline()}: assigns an
    \texttt{OfflineMultiplayerPeer} so the game runs in single-player
    with no network. The local instance is the authority for all nodes;
    \texttt{is\_multiplayer\_authority()} is \texttt{true} where you
    expect, and no packets are sent.
\end{itemize}

\paragraph*{Queries}
\begin{itemize}
  \item \texttt{is\_server()}: \texttt{true} if this instance is the
    server (or host).
  \item \texttt{is\_offline()}: \texttt{true} if
    \texttt{OfflineMultiplayerPeer} is in use.
  \item \texttt{get\_local\_id()}: unique peer ID of this instance (e.g.\
    for UI or logging).
\end{itemize}

\paragraph*{Signals}
\begin{itemize}
  \item \texttt{peer\_connected(id, name)}, \texttt{peer\_disconnected(id,
    name)}: when a peer joins or leaves (server-side).
  \item \texttt{connection\_succeeded}, \texttt{connection\_failed}:
    client-side connection result.
  \item \texttt{server\_started}: after \texttt{host\_game()} succeeds.
\end{itemize}

\paragraph*{RPCs (examples)}
\begin{itemize}
  \item \texttt{send\_message\_to\_server(text)}: \texttt{@rpc("any\_peer")};
    clients call this to send a string to the server; the server prints
    the sender ID and the text.
  \item \texttt{broadcast\_message(text)}: \texttt{@rpc("any\_peer",
    "call\_local")}; runs on all peers including the caller, so the host
    sees it too. Useful for chat or global announcements.
\end{itemize}

\subsection*{How to use it in game code}

\paragraph*{From any node}
After the autoload is set up, call the API from menus or game logic:
\begin{verbatim}
NetworkManager.host_game(8080)
NetworkManager.join_game("127.0.0.1", 8080)
NetworkManager.start_offline()
NetworkManager.leave_game()
\end{verbatim}

\paragraph*{Authority and offline}
Use \texttt{multiplayer.is\_multiplayer\_authority()} (or the node's
\texttt{is\_multiplayer\_authority()}) so only the owner updates that
node. In offline mode the local peer is authority everywhere, so the
same logic runs without branching:
\begin{verbatim}
if multiplayer.is_multiplayer_authority():
    move_character(delta)
\end{verbatim}

\paragraph*{Calling RPCs}
The node calling the RPC must be in the scene tree (e.g.\ under the
same root as \texttt{NetworkManager}). Use \texttt{.rpc()} for the
current peer or \texttt{.rpc\_id(id)} for a specific peer:
\begin{verbatim}
NetworkManager.send_message_to_server.rpc("Hello from client")
NetworkManager.broadcast_message.rpc("Hello everyone")
\end{verbatim}

\paragraph*{Using the menu scene}
Add \texttt{multiplayer\_menu.tscn} as a child of the main scene, or run
it as a test scene. Buttons drive \texttt{NetworkManager}; optional
\texttt{IPEdit} and \texttt{PortEdit} override the default
\texttt{127.0.0.1} and \texttt{8080}. The status label shows connection
state. For single-player, use ``Offline / Single Player'' so the game
uses \texttt{OfflineMultiplayerPeer} and no network port is opened.

\end{document}
